Hou10ni Profiling with TAU

### 2017-05-04 Paper2/answer-reviewers.tex

We thank the  reviewers for their interest in our work and for their helpful comments that will greatly improve  the manuscript. We have  tried  to do  our  best to  respond to  the  points  raised. The reviewers have brought  up  some  good  points  and  we  appreciate  the  opportunity  to  clarify  our research objectives and results. As  indicated  below,  we  have  checked  all  the  general  and  specific  comments  provided  by  the reviewers and have made necessary changes accordingly to their indications.\\

\noindent{\large\bf Anonymous reviewers \#1}\\

\noindent
{\em 1) The usage workflow is not that clear. For instance how optimisations can be selected ? Is there a list of such optimisations or is it the user
  to design the optimisation sets. In Section V.A how different values are selected? Is there any way for the framework to guide the user towards
  optimisations or to give feedback on what is working, what is not working, etc. The authors mention MAQAO in the conclusion... Overall, this part is
  really fuzzy.}\\

The workflow has been better described (see diff file attached).\\

@@

\noindent
{\em 2) Listing 3 should be explained in details or removed. Not every reader is a specialist of vectorisation}\\

Listing 3 and vectorization have been described carefully in Section B.a.\\

@@

\noindent
{\em 3) section V.A page 8. You state that you have an unexpected results that should be investigate. A journal paper is not a conference paper. I
  think you should investigate and publish what you have found in ***this*** paper.} \\

Autotuning is useful because it can challenge our assumptions. We modified a sentence accordingly. \\

@@

\noindent
{\em 4) Section V.C. is not easy to follow. It is hard to understand where the problems you faced came from : from the kernels, from the design of BOAST, from a bug in the implementation, etc. What kind of conclusion should be taken from your experience in the usage of BOAST? Such discussion should help the reader in doing better science. The writing of this section does not go in that direction. What can be generalised from your experience? what are your advices?}\\

Our approach's objectives have been better described at the beginning of the section. Then several paragraphs of this section have been modified to
better emphasize the results and expectations of this part of our work.\\

@@

\noindent
{\em 5) in Section V.C why did you not generate the CUDA version as well and compare it with the hand-tuned existing version? For the revised version, I need to see that experiment added to paper!}\\

Indeed this experience was already performed. We rewrote this section to better mention our results around CUDA.

### 2017-06-14 hou10ni/Tutorial/hou10ni_profile.tex
Hou10ni's kernel : Optimisation

\begin{cverbatim}
git clone https://github.com/Nanosim-LIG/boast.git
cd boast
gem build BOAST.gemspec
gem install --user-install BOAST-1.3.5.gem 
\end{cverbatim}

@@

We generate the exact same kernel as a reference version and run it with BOAST in order to compare the results in term of correctness and performances. KRef defines the reference version. This version simply need to keep the Fortran code (referred as \texttt{codeF}) and wrap it with BOAST.

@@

The file run.rb is used to launch the kernel.

\begin{lstlisting}[caption=run.rb]
require 'BOAST'
require './KRef.rb'
require 'narray_ffi'

@@

\begin{lstlisting}[caption=Results (obtained with the command ruby run.rb)]
SUBROUTINE fluid_inner_elt_ref(Nflu_inner, Nflusol_inner, nb_rhs, idx_&
&vec_flu, idx_mat_flu, P_new, P_inter, P_old, A_flu, afluSize, pSize, v&
&ecSize_1, vecSize_2, matSize)
  implicit none
  INTEGER, PARAMETER :: dq=8 !< quadruple precision
  INTEGER, PARAMETER :: dp=4  !< double precision

@@

TODO

>>>>

\begin{cverbatim}
git clone https://github.com/Nanosim-LIG/boast.git
cd boast
gem build BOAST.gemspec
gem install --user-install BOAST-2.0.3.gem 
\end{cverbatim}

@@

We generate the exact same kernel as a reference version and run it with BOAST in order to compare the results in term of correctness and performances. KRef defines the reference version. This version simply needs to keep the Fortran code (referred as \texttt{codeF}) and wrap it with BOAST. To put it in a nutshell, a reference version should follow the template bellow.

\lstset{language=RUBY}
\begin{lstlisting}[caption=KRef\_template.rb]
def KRef
  lang = BOAST::get_lang
  BOAST::set_lang(BOAST::FORTRAN)
  kernel = BOAST::CKernel::new
  function_name = "function_ref"

  VARIABLES HERE
  FOR EXAMPLE : 
  geocode = BOAST::Int("geocode", :dir => :in )
  n01 = BOAST::Int("n01", :dir => :in)
  u = BOAST::Real("u", :dir => :in, :dim => [ BOAST::Dim(0, n01),  BOAST::Dim(0, n02))] )

   p = BOAST::Procedure::new(function_name, [geocode,n01,u])

   kernel.code.print <<EOF
  ORIGINAL C or FORTRAN CODE HERE
EOF

  kernel.procedure = p
  BOAST::set_lang(lang)
  return kernel
end

geocode = ...
n01 = ...
u = ...
k = KRef
stats = k.run(geocode, n01, u)
puts "#{k.procedure.name}: #{stats[:duration]*1.0e3}"
\end{lstlisting}

We wrote the reference version a bite differently as shown bellow.

@@

The file run\_ref.rb is used to launch the kernel.

\begin{lstlisting}[caption=run\_ref.rb]
require 'BOAST'
require './KRef.rb'
require 'narray_ffi'

@@

\begin{lstlisting}[caption=Results (obtained with the command ruby run\_ref.rb)]
SUBROUTINE fluid_inner_elt_ref(Nflu_inner, Nflusol_inner, nb_rhs, idx_&
&vec_flu, idx_mat_flu, P_new, P_inter, P_old, A_flu, afluSize, pSize, v&
&ecSize_1, vecSize_2, matSize)
  implicit none
  INTEGER, PARAMETER :: dq=8 !< quadruple precision
  INTEGER, PARAMETER :: dp=4  !< double precision

@@

\lstset{language=RUBY}
\begin{lstlisting}[caption=KBoast.rb]
class KBoast
 attr_reader :kernel

 def initialize(options)

  @opts = {:preprocessor => false, :unroll => false, :inline => :included}
  @opts.update(options)

  # Parameters 
  @Nflu_inner         = Int("Nflu_inner", :dir => :in )
  @Nflusol_inner      = Int("Nflusol_inner", :dir => :in )
  @nb_rhs             = Int("nb_rhs", :dir => :in )
  @afluSize           = Int("afluSize", :dir => :in )
  @pSize              = Int("pSize", :dir => :in )
  @vecSize_1          = Int("vecSize_1", :dir => :in )
  @vecSize_2          = Int("vecSize_2", :dir => :in )
  @matSize            = Int("matSize", :dir => :in )

  @idx_vec_flu        = Int("idx_vec_flu", :dir => :in, :dim => [Dim(@vecSize_1),Dim(@vecSize_2)] )
  @idx_mat_flu        = Int("idx_mat_flu", :dir => :in, :dim => [Dim(@matSize)] )
  @P_new              = Real("P_new",:dir => :inout, :dim => [Dim(@pSize),Dim(@nb_rhs)])
  @P_inter            = Real("P_inter",:dir => :in, :dim => [Dim(@pSize),Dim(@nb_rhs)])
  @P_old              = Real("P_old",:dir => :in, :dim => [Dim(@pSize),Dim(@nb_rhs)])
  @A_flu              = Real("A_flu",:dir => :in, :dim => [Dim(@afluSize)])

 end

 def generate

   push_env(:lang => FORTRAN) {

    @kernel = CKernel:: new()
    @kernel.procedure = Procedure("fluid_inner_elt_boast",[@Nflu_inner, @Nflusol_inner, @nb_rhs, @idx_vec_flu, @idx_mat_flu, @P_new, @P_inter, @P_old, @A_flu, @afluSize, @pSize,@vecSize_1,@vecSize_2,@matSize] ,:functions => nil)

    #..  FUNCTIONS ..#
    register_funccall("matmul") if get_lang == FORTRAN
    call_matmul = lambda{|x,y|
      if get_lang == FORTRAN
        return matmul(x,y)
      end
    }

    #.. KERNEL ..#

    opn @kernel.procedure

    # local variables
    i = Variable::new('I', Int)
    j = Variable::new('J', Int)
    i_tmp1 = Variable::new('I_tmp1', Int)
    i_tmp2 = Variable::new('I_tmp2', Int)
    i1 = Variable::new('I1', Int)
    i2 = Variable::new('I2', Int)
    i3 = Variable::new('I3', Int)
    i4 = Variable::new('I4', Int)
    p_aux = Variable::new("P_aux",Real, :dimension => [Dim(20),Dim(@nb_rhs)])

    decl i,j,i_tmp1,i_tmp2, i1,i2,i3,i4,p_aux

    pr For(i,1,@Nflu_inner+@Nflusol_inner){
      pr i_tmp1 === 1
      pr For(j,1,@idx_vec_flu[1,i]){
        pr i1 === @idx_vec_flu[2*j,i]
        pr i2 === @idx_vec_flu[2*j+1,i]
        pr i_tmp2 === i_tmp1 + i2 - i1
        pr p_aux.slice(i_tmp1..i_tmp2,1..@nb_rhs) === @P_inter.slice(i1..i2,1..@nb_rhs) 
        pr i_tmp1 === i_tmp2 +1
      }
      pr i1 === @idx_vec_flu[2,i]
      pr i2 === @idx_vec_flu[3,i]
      pr i3 === @idx_mat_flu[i]
      pr i4 === @idx_mat_flu[i+1] - 1
      reshape_code ="RESHAPE(#{@A_flu.slice(i3..i4)}, (/#{i2}-#{i1}+1, #{i_tmp2}/))"
      pr @P_new.slice(i1..i2,1..@nb_rhs) === call_matmul.call(reshape_code , p_aux.slice(1..i_tmp2,1..@nb_rhs) ) - @P_old.slice(i1..i2,1..@nb_rhs)
    }
    close @kernel.procedure
   }

   return @kernel
 end
end
\end{lstlisting}

\noindent The file run\_boast.rb is used to launch the BOAST version of the kernel. We run 3 times the kernel (repeat). \\

\begin{lstlisting}[caption=run\_boast.rb]
require 'BOAST'
require './KBoast.rb'
require 'narray_ffi'

 def self.run()

 # set_default_int_size(8)
  set_default_real_size(4)
  k_boast_params = {:kernel => :boast, :preprocessor => false, :LDFLAGS => "-lgfortran", :FCFLAGS => "-fbounds-check -fimplicit-none"}

  set_lang(FORTRAN)
  kernels = {}
  stats_svg = {}
  stats = []
  repeat = 3

  # Creating ref kernel
  kernels[k_boast_params] = KBoast::new(k_boast_params)
  kernels[k_boast_params].generate
  puts "*********** Building kernel ***********\n"
  puts kernels[k_boast_params].kernel
  kernels[k_boast_params].kernel.build(:LDFLAGS => k_boast_params[:LDFLAGS], :FCFLAGS => k_boast_params[:FCFLAGS] )

  puts "\n*********** Testing boast kernel with inputs parameters ***********\n"
  inputs = kernels[k_boast_params].kernel.load_ref_inputs()
  outputs = kernels[k_boast_params].kernel.load_ref_outputs()

  repeat.times {
    inputs.each_key { |key|
      puts key
      stats.push kernels[k_boast_params].kernel.run(*(inputs[key]))
      puts kernels[k_boast_params].kernel.run(*(inputs[key])).inspect
      puts kernels[k_boast_params].kernel.compare_ref(outputs[key], inputs[key]).inspect
    }
  }
  stats.sort_by! { |a| a[:duration] }
  stats = stats.first

  puts "#{kernels[k_boast_params].kernel.procedure.name}: #{stats[:duration]} s ->  #{stats[:duration]*1.0e3} ms"
  return kernels
 end
end

\begin{lstlisting}[caption=Results (obtained with the command ruby run\_boast.rb)]
SUBROUTINE fluid_inner_elt_boast(Nflu_inner, Nflusol_inner, nb_rhs, idx_&
&vec_flu, idx_mat_flu, P_new, P_inter, P_old, A_flu, afluSize, pSize, v&
&ecSize_1, vecSize_2, matSize)

  integer, parameter :: wp=kind(1.0d0)
  integer(kind=4), intent(in) :: Nflu_inner
  integer(kind=4), intent(in) :: Nflusol_inner
  integer(kind=4), intent(in) :: nb_rhs
  ...

END SUBROUTINE fluid_inner_elt_boast

*********** Testing boast kernel with inputs parameters ***********
./fluid_inner_elt_boast/sub_time_loop_1
{:duration=>0.13263683, :start=>-4004705691753970176, :end=>-3872068861753970176}
{"P_new"=>0.0}
./fluid_inner_elt_boast/sub_time_loop_1
{:duration=>0.140625219, :start=>-3642235174753970176, :end=>-3501609955753970176}
{"P_new"=>0.0}
./fluid_inner_elt_boast/sub_time_loop_1
{:duration=>0.147113606, :start=>-3274643277753970176, :end=>-3127529671753970176}
{"P_new"=>0.0}
fluid_inner_elt_boast: 0.15179012 s ->  151.79012 ms  
\end{lstlisting}

\subsubsection*{Optimisations}
